Transformers, do we really need them for detecting stress?

### 2023-06-04 paper/paper.tex
Close to finishing paper

\usepackage[utf8]{inputenc}
\usepackage[pdftex]{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{cleveref}

\title{Stress Analysis in Social Media}

@@

This section is the introduction to your paper.

\section{Related Work}

In scientific papers, this section usually (but not necessarily) briefly describes the related research and what makes the presented approach different from it.

@@

\section{Models}
Motivated by the results of the before mentioned paper, our approach focused on three different models which solve the given classification problem.
\hfill \break
\hfill \break
Firstly we will introduce the \textbf{logistic regression} model which the authors of the dataset presented as the best solution, excluding state-of-the-art solutions such as BERT.
Then we are going to introduce two transformer approaches, one being \textbf{DistilBERT} and the other one, currently very popular, \textbf{Chat GPT 3.5 Turbo}.

@@

\subsection{Logistic regression}
Logistic regression is a commonly used machine learning model which can be used for binary classification problem.
This is a simple model which doesn't require long and expensive training, so, inspired by the great results that this model showed in the original paper (F-score of 79.8 ~\citep{turcan-mckeown-2019-dreaddit}), we have also decided to
include it in our own research.
\hfill \break
\hfill \break
Logistic regression model expects a fixed size input but words and sentences in natural language can be of any size.
Today this problem is solved by using word embeddings which represent words from our dataset as multidimensional vectors. Their advantage over some other approaches that were used in the past, such as one-hot vector representation, is that they are able to find relations between words and produce similar vectors for similar words.In our approach we used \textbf{word2Vec}~\citep{mikolov2013efficient} embeddings which were trained on the whole training dataset and produced 300-dimensional representations.
Learned word embeddings were then used for calculating the final representation for each training example.
Here we experimented with 2 approaches:
\begin{enumerate}
    \item \label{itm:one} \textbf{Averaging of word embeddings:} summing up each representation in the input sequence and dividing it with the number of words in the sequence
    \item \textbf{TF-IDF averaging:} Term Frequency - Inverse Document Frequency is an algorithm that is used to determine an importance of the given word in the document.
    Importance is represented as a scalar value.
     This scalar value for each word is then multiplied with the word's corresponding representation.
    The final representation is then calculated in the same way as described in~\ref{itm:one}.

@@

\subsection{Chat GPT 3.5 Turbo}

@@

The final model that we used in our research was Chat GPT 3.5 Turbo.
This model was developed by OpenAI and it aims to generate human like responses based on the provided input text.
We couldn't perform direct training on this model, but instead we used the provided API to prompt the model to give us the classification result for each example in the test set.
You can see how we formulated our prompt in figure~\ref{fig:chat-gpt-prompt}.

Given results were then cleaned up

@@

Finally, a section which describes the acquired results.

\section{Conclusion}

Conclusion is the last enumerated section of the paper.
It should not exceed half of a column and is typically split into 2--3 paragraphs.
No new information should be presented in the conclusion; this section only summarizes and concludes the paper.

>>>>

\usepackage[utf8]{inputenc}
\usepackage[pdftex]{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{cleveref}
\usepackage{biblatex}
\addbibresource{}

\title{Transformers, do we really need them for detecting stress?}

@@

Stress is a very common feeling everyone experiences in a smaller or larger amount during their life.
It negatively impacts our health and detecting it could enable us to help the person who is experiencing it and improve their existing situation.
This paper will try to find the optimal approach of detecting stressful situations from user descriptions.
For this task we will be using the Dreaddit dataset which contains user posts from the Reddit website.
Throughout this paper we will be referencing the original one where the dataset was introduced.
\hfill \break
\hfill \break
In the first part of the paper we are going to describe the dataset, followed by the models that we decided to use to tackle the problem.
We tried to continue the research of the authors of the dataset and use their best performing model to compare it with some other currently popular solutions.
In the end we are going to try to answer the question if the more complex and costly solutions were really justified or should this problem be solved with a more simple approach.

\section{Related Work}
Stress detection was researched in a couple of other approaches in the last couple of years.
One of the approaches was detecting stress using deep neural networks and psychological signals~\citep{stress-neural}.

The approach that we looked at the most was by the authors of the dataset\

@@

\section{Models}
Motivated by the results of the aforementioned paper, our approach focused on three different models which solve the given classification problem.
\hfill \break
\hfill \break
Firstly we will introduce the \textbf{logistic regression} model which the authors of the dataset presented as the best solution, excluding state-of-the-art (SOTA) solutions such as BERT.
This model is going to represent a simple model approach to the problem without using more complex solutions such as transformers.
Then we are going to introduce two transformer approaches, one being \textbf{DistilBERT} and the other one, currently very popular, \textbf{Chat GPT 3.5 Turbo}.

We performed our training on the whole dataset and our focus wasn't to get the best possible results for each model but to compare them and find out which is best suited for the given classification task (taking into account the cost of the model).
Because of that reason we also didn't remove lower agreement rate examples from our training dataset to see how our models deal with learning harder examples.

@@

\subsection{Logistic regression}
Logistic regression~\citep{logreg} is a commonly used machine learning model which can be used for binary classification problem.
This is a simple model which doesn't require long and expensive training, so, inspired by the great results that this model showed in the original paper (F-score of 79.8 ~\citep{turcan-mckeown-2019-dreaddit}), we have also decided to
include it in our own research.
\hfill \break
\hfill \break
Logistic regression model expects a fixed size input but words and sentences in natural language can be of any size.
Today this problem is solved by using word embeddings which represent words from our dataset as multidimensional vectors. Their advantage over some other approaches that were used in the past, such as one-hot vector representation, is that they are able to find relations between words and produce similar vectors for similar words.In our approach we used \textbf{word2Vec}~\citep{mikolov2013efficient} embeddings which were trained on the whole training dataset and produced 300-dimensional representations.
Learned word embeddings were then used for calculating the final representation for each training example.
Here we experimented with 2 approaches:
\begin{enumerate}
    \item \label{itm:one} \textbf{Averaging of word embeddings:} summing up each representation in the input sequence and dividing it with the number of words in the sequence
    \item \textbf{TF-IDF averaging:} Term Frequency - Inverse Document Frequency is an algorithm that is used to determine an importance of the given word in the document.
    Importance is represented as a scalar value.
     This scalar value for each word is then multiplied with the word's corresponding representation.
    The final representation is then calculated in the same way as described in~\ref{itm:one}.

@@

@@

\subsection{Chat GPT 3.5 Turbo}
The final model that we used in our research was Chat GPT 3.5 Turbo.
This model was developed by OpenAI and it aims to generate human like responses based on the provided input text.
We couldn't perform direct training on this model, but instead we used the provided API to prompt the model to give us the classification result for each example in the test set.
You can see how we formulated our prompt in figure~\ref{fig:chat-gpt-prompt}.

It must be mentioned that there was some manual data cleanup needed after we got the responses from the API (labels contained an explanation behind the given result)

\begin{table*}
    \centering
    \caption{Different model results}
    \label{tab:model-results}
    \begin{center}
        \begin{tabular}{|l|c|c|c|}
            \hline
            \textbf{Model} & \textbf{P} & \textbf{R} & \textbf{F1} \\ \hline
            LR + averaged word2Vec embeddings & 0.6441 & 0.7995 & 0.7134 \\ \hline
            LR + TF-IDF averaged word2Vec embeddings & 0.6859 & 0.8049 & \textbf{0.7406} \\ \hline
            DistilBERT & 0.7105 & 0.7913 & 0.7487 \\ \hline
            Chat GPT 3.5 Turbo & 0.6905 & \textbf{0.9431} & 0.7972 \\ \hline
        \end{tabular}
    \end{center}
\end{table*}

@@

The results of our four different models are present in table~\ref{tab:model-results}.
We can see that our logistic regression (LR) model with basic word embedding averaging performed the worst, what was expected given its simplicity.
This simple model, however, performs much better when we introduce TF-IDF averaging.
\subsection{LR(TF-IDF averaging) vs DistilBERT}
LR model with TF-IDF averaging performed very similarly to the DistilBERT approach.
Transformers are currently SOTA approach in NLP and we expect them to perform much better than the simpler models.
However, classifying stress, using this pretty clean dataset, isn't a particularly hard task and our logistic regression model deals very well with the problem, which is also what authors of the original dataset pointed out.
\hfill \break
\hfill \break
Here we compared only the results of the two models not taking into account the cost of the more complex model.
Training of the DistilBERT model took longer and required more hardware just to get worse results.
This leads us to the conclusion that sometimes the solution to the problem could be a simple model without the need for more complex and expensive solutions.
\\

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{images/conf-matrix-chatgpt}
    \caption{Confusion matrix for Chat GPT 3.5 Turbo. 0 represents data labeled not stressed while 1 represents data labeled stressed.}
    \label{fig:chat-gpt-conf-matrix}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{images/conf-matrix-logreg}
    \caption{Confusion matrix for LR with TF-IDF averaging.}
    \label{fig:logreg-conf-matrix}
\end{figure}

\subsection{DistilBERT vs Chat GPT}
In our two transformer approaches Chat GPT showed much better results.
What shows the true power of this transformer model is that, even though it wasn't strictly trained on the training data, it managed to outperform DistilBERT model.
As we can see in figure ~\ref{fig:chat-gpt-conf-matrix} this model managed to have an extremely low number of false negative (FN) predictions.
In certain cases we would want to prefer this behaviour where we want to detect stress and help the person who is experiencing it.

The downside of this approach is the cost behind it.
Chat GPT API is a paid service and if we want to use this approach in for stress detection this might not be the best option if our budget is limited.

\subsection{LR (TF-IDF) vs Chat GPT}
We can see that Chat GPT performed better than our LR model when we compare their F1 scores.
The change comes primarily from the higher recall score which Chat GPT manages to achieve (less false negative examples).
While we can look at the scores by themselves and conclude that Chat GPT is a better model for the given task, this approach doesn't take into account the cost of the model.

Our simpler LR model achieves comparable results to the SOTA solution and we claim that it is the optimal approach for solving stress classification problem from the given dataset.
Model is easy to learn and could even get better results if it was trained on the dataset which contained examples with higher agreement factor, as it was shown from the authors of the dataset.
\section{Conclusion}
